\documentclass[letterpaper,10pt,onecolumn]{IEEEtran}

\usepackage{geometry}
\usepackage{graphicx}
\usepackage{caption}

\usepackage[utf8]{inputenc}
\geometry{margin=0.75in}


\title{Project 4: The SLOB SLAB}
\author{Christian Armatas and Mihai Dan}
\date{November 9th, 2016}

\begin{document}

    \begin{center}
        \begin{minipage}[h]{\textwidth}
            \maketitle
        \end{minipage}
    \end{center}
    
    \vspace{140mm}
    
    \begin{center}
        \section*{Abstract}
        This document contains the design for necessary SLOB first-fit algorithms, a version control log table, a work log, and answers to the provided questions. The solution to Project 4 can be deduced from the previously mentioned sections, as well as the implementation. The purpose of this project was to better our understanding of *not sure yet*
    \end{center}
    
    
    \newpage
    
    
    \section*{Design for Implementation of Necessary Algorithms}
  

    \vspace{6mm}
    
    
    \section*{Version Control Log}
        \begin{center}
            Table organized by most recent commit.
        \end{center}
        
        \vspace{0.5mm}
        
        \begin{center}
        \def\arraystretch{1.1}
        \begin{tabular}{ | p{8cm} | }
            \hline
            
            \hline
        \end{tabular}
        \end{center}


    
    \vspace{6mm}
    
    
    \section*{Work Log}
    
    
    \begin{itemize}
        \item Day, 11/??:
            \begin{itemize} 
                \item 
        	\end{itemize}
        \item Day, 11/??:
            \begin{itemize} 
                \item 
        	\end{itemize}
        \item Day, 11/??:
            \begin{itemize} 
                \item 
        	\end{itemize}
    \end{itemize}
    
    
    \vspace{6mm}
    
   
    \section*{Questions}
    \begin{enumerate}
        \item What do you think the main point of this assignment is?
        \begin{itemize}
            \item Answer.
        \end{itemize}
        \item How did you personally approach the problem? Design decisions, algorithm, etc.
        \begin{itemize}
            \item Answer
        \end{itemize}
        \item How did you ensure your solution was correct? Testing details, for instance.
        \begin{itemize}
            \item Answer.
        \end{itemize}
        \item What did you learn?
        \begin{itemize}
            \item Answer.
        \end{itemize}
    \end{enumerate}
    
    
\end{document}
